---
title: "A look at NYC Shooting Data from 2020"
author: "Samir Fidai"
date: "2024-04-25"
slug: []
categories: []
tags: []
---

For my first blog post in well over a year, I chose to take a further look at the NYPD's publicly available Shooting Incident Data. The link to the public dataset used for this blog post can be found [here](https://data.cityofnewyork.us/Public-Safety/NYPD-Shooting-Incident-Data-Historic-/833y-fsy8/about_data). As always, all the code that I have written to create the visualizations below will be available to view on my [Github](https://github.com/FidaiGuy/NYC-mini-blog).




```{r setup, include=FALSE}
library(tidyverse)
library(plotly)
library(leaflet)
library(ggpubr)
library(mapview)

theme_set(theme_minimal())
```

```{r, include=FALSE}
NYPD_shooting_data <- read_csv("C:/Users/samirfidai/Downloads/NYPD_Shooting_Incident_Data__Historic__20240422.csv")
NYPD_shooting_data %>% glimpse()
```

```{r, include=FALSE}
NYPD_shooting_data <- NYPD_shooting_data %>%
  mutate(
    OCCUR_DATE = mdy(OCCUR_DATE),
    OCCUR_YEAR = year(OCCUR_DATE)
  )
```




The visualization below is a trended look at the number of Shootings by year. I've broken it out by borough as well as the race of the shooting victim. We can see a majority of recorded shooting incidents occur in The Bronx and Brooklyn with African Americans being the primary victim. What really stood out to me was the massive spike between 2019-2020. I decided to focus the rest of this blog in doing further investigations on *when* these shootings were occurring and *where* in the boroughs were shooting incidents concentrated. 

```{r, fig.width=14, fig.height = 8, echo=FALSE, warning=FALSE,results="hide"}

murders_by_borough <- NYPD_shooting_data %>%
  mutate(occur_year = year(OCCUR_DATE)) %>%
  group_by(BORO, occur_year) %>%
  summarize(murders = n()) %>%
  ggplot(aes(occur_year, murders)) +
    geom_line(aes(color = BORO)) +
    scale_x_continuous(breaks = seq(2006, 2024, 1)) +
    labs(
      title = "Shooting Incidents By Borough",
      subtitle = "Note: 2024 data is through March 2024.",
      x="year")+
    theme(legend.position = "top")

murders_by_race <- NYPD_shooting_data %>%
  mutate(occur_year = year(OCCUR_DATE)) %>%
  group_by(VIC_RACE, occur_year) %>%
  summarize(murders = n()) %>%
  ggplot(aes(occur_year, murders)) +
    geom_line(aes(color = VIC_RACE)) +
    scale_x_continuous(breaks = seq(2006, 2024, 1)) +
    labs(
      title = "Shooting Incidents By Race",
      subtitle = "Note: 2024 data is through Q1 March",
      x = "year"
    ) +
    theme(legend.position = "top")

ggpubr::ggarrange(murders_by_borough, murders_by_race, ncol = 1, widths = c(9,1))
```

<br>
<br>
<br>

To visualize the time when shootings were occurring in NYC by borough, I created a heatmap in ggplot. In most boroughs, shootings occur primarily in the evening/night and rarely during the day. This pattern can be seen in Brooklyn where there were 355 reported shootings in 2020 that occurred between 9pm and 1am. 

```{r, fig.width=10, echo=FALSE}
ggplotly(
NYPD_shooting_data %>%
  mutate(OCCUR_HOUR = hour(OCCUR_TIME),) %>%
  dplyr::filter(OCCUR_YEAR >= 2020 & OCCUR_YEAR < 2021) %>%
  group_by(BORO, OCCUR_HOUR) %>%
  summarize(n = n()) %>%
  ggplot(aes(x = OCCUR_HOUR, y = BORO, fill = n))+
    geom_tile() +
    scale_fill_gradient(low = "lightblue", high = "darkblue", na.value = NA) +
    labs(title = "Time of Shooting Incident in NYC in 2020 by Borough"))
```



<br>
<br>
<br>

Finally, I mapped the location of each shooting in NYC in 2020. Using the included geographical coordinates provided in the Data, I used the Leaflet Package in R to create a map of shooting incidents. The NYPD flags shootings if they were counted as a murder. When you click a shooting incident, a user would be able to view the incident date and demographics on the victim.

**Location of NYC Shooting Incidents in 2020**
```{r, fig.width=10, echo=FALSE}
NYPD_shooting_data_2021 <- NYPD_shooting_data %>%
  dplyr::filter(OCCUR_YEAR >= 2020 & OCCUR_YEAR < 2021) 

popup_view = paste0(
  "<b> Incident Date: </b>",
  NYPD_shooting_data_2021$OCCUR_DATE,
  "<br>",
  "<b>Victim's Age: </b>",
  NYPD_shooting_data_2021$VIC_AGE_GROUP,
  "<br>",
  "<b>Victim's Sex: </b>",
  NYPD_shooting_data_2021$VIC_SEX,
  "<br>",
  "<b>Victim's Race: </b>",
  NYPD_shooting_data_2021$VIC_RACE
)


NYPD_shooting_data_2021 %>% leaflet() %>%
  addTiles() %>%
  addCircleMarkers(~NYPD_shooting_data_2021$Longitude, 
                   ~NYPD_shooting_data_2021$Latitude,
                    clusterOptions = markerClusterOptions(),
                   color = ~ifelse(STATISTICAL_MURDER_FLAG == TRUE, "red", "blue"),
                   popup = ~popup_view) %>%
  setView(-74.00, 40.71, zoom = 11) %>%
  addLegend("bottomleft",
            colors = c("red", "blue"),
            labels = c("Yes", "No"),
            title = "Shooting counted as a murder")
```


